### 2024-09-23 paper/paper.qmd
Draft of the report
1. upload a draft of report.
2. some references added.

---
title: "Rising Trends in Shelter Resident Deaths in Toronto Over the Past 17 Years"
author: 
  - Haowei Fan
thanks: "Data are available at: https://open.toronto.ca/dataset/deaths-of-shelter-residents/"
date: today
date-format: long
abstract: "Shelter resident deaths are an important indicator of social stability, reflecting systemic issues such as welfare systems, housing crises, and healthcare access. This project analyzed monthly shelter resident deaths in the Toronto region from March 2007 to August 2024, focusing on trends over time and seasonal variations. The study found that deaths have been increasing, with a significant rise between 2020 and 2023. While the number of deaths remained similar in spring, summer, and fall, winter months saw a significantly higher number of fatalities. These findings provide valuable data to support targeted interventions by governments and welfare organizations to address changing needs throughout the year."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false
#| echo: false
library(tidyverse)
library(knitr)
analysis_data <- read_csv("/Users/fhw0912/Desktop/Deaths-of-Shelter-Residents-in-Toronto/data/analysis_data/analysis_data.csv")
```

# Introduction

In 2023, as the COVID-19 pandemic receded, the Toronto government announced plans to phase out the shelter hotels that had been providing temporary accommodations for homeless individuals. By the end of 2024, 24 shelter hotels will no longer be available due to expiring contracts, exacerbating an already overcrowded shelter system. The loss of these temporary accommodations poses significant risks not only to the safety and stability of the community but also to the well-being of those who rely on shelters for survival. The influx of new shelter residents may potentially lead to a decline in living conditions, including sanitation issues and territorial conflicts. This highlights the urgent need to understand the trends and underlying factors affecting shelter residents, especially in terms of their mortality numbers.
While many sociological studies have explored the growing number of shelter residents, fewer have examined shelter resident mortality rates in detail. Given the challenges of accurately counting the transient and often untracked shelter population, studying actual death counts offers a more reliable and feasible method of understanding the impact of systemic issues. Mortality data, being both concrete and verifiable, can better inform policy decisions than estimates of shelter population numbers, which are often difficult to obtain and track.
This paper focuses on analyzing the number of deaths among shelter residents in Toronto from March 2007 to August 2024. The aim is to identify both long-term trends and seasonal variations in shelter resident deaths. Using simple linear regression, this study estimates the rate at which the number of deaths has increased over time. It also examines anomalies in the data by comparing yearly death rates, and assesses seasonal differences by analyzing average death rates across the four seasons. This analysis provides valuable insights into the hardships faced by shelter residents, especially during specific times of the year.
This study finds three key trends: 1) the number of deaths among shelter residents has been rising over time; 2) there was a noticeable increase in deaths between 2020 and 2023; and 3) while deaths during spring, summer, and fall remained relatively consistent, winter saw significantly higher mortality rates. Understanding these trends is critical because shelter resident deaths serve as a key indicator of broader social problems, such as gaps in welfare systems, housing shortages, and insufficient healthcare services. These findings provide essential data for governments and welfare organizations to develop more targeted interventions based on the needs of shelter residents throughout the year.
The remainder of this paper is structured as follows: Section 2 discusses the data sources used in this project and details the data cleaning process, followed by a visualization of the dataset's key distribution and statistics. Section 3 outlines the statistical methods used and presents the findings in detail. Section 4 addresses the limitations of the study and offers suggestions for future research.

# Data {#sec-data}

## Raw Data

The dataset I am using is titled "Deaths of Shelter Residents", sourced from OpenToronto. This dataset records the number of deaths among shelter residents in the Toronto area, starting from January 2007 on a monthly basis. It also includes the gender of the deceased, categorized into three sections: male, female, and other gender, which includes Transgender, Non-binary, and Two-Spirit. The dataset is continuously updated. Table 1 contains descriptions of these variables.

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: Column descriptions of the original data
#| tbl-cap: "Column descriptions of the original data"

column_description <- tibble(
  Column = c("_id", "Year", "Month", "Total decedents", "Male", "Female", "Transgender/Non-binary/Two-Spirit"),
  Description = c(
    "Unique row identifier for Open Data database.",
    "The calendar year being reported on.",
    "The month being reported on.",
    "The total number of shelter residents who died in the reported month/year.",
    "The total number of male shelter residents who died in the reported month/year.",
    "The total number of female shelter residents who died in the reported month/year.",
    "The total number of transgender, non-binary, and Two-Spirit shelter residents who died in the reported month/year."
  )
)

kable(column_description, col.names = c("Column", "Description"))
```

In addition to this dataset, OpenToronto also offers a similar dataset called "Deaths of People Experiencing Homelessness", which includes four variables: id, year, month, and number of deaths. However, I chose not to use this dataset due to its limited data coverage, spanning only from May 2007 to August 2023. In comparison, the "Deaths of Shelter Residents" dataset provides more extensive data, allowing for greater accuracy and a narrower confidence interval in the results.

## Data Cleaning

In the data cleaning process, the following steps were conducted using Rcode(@citeR), along with the packages pandas, tidyverse, ggplot2, and dplyr:

Combining Year and Month: The variables for year and month were merged, and the month was converted to a numerical format to facilitate time series analysis.
Creating a New Variable for Seasons: A new variable was created to categorize each month into one of the four seasons: March, April, and May were classified as spring; June, July, and August as summer; September, October, and November as autumn; and December, January, and February as winter. This was done to enable seasonal analysis.
Removing Unnecessary Variables: All variables unrelated to this project, such as id and the three gender categories, were removed to simplify the dataset.
Ensuring Equal Seasonal Data: To ensure that each season had the same length of time for comparison, I removed some samples from the beginning and end of the dataset, so that the total number of samples became a multiple of four, with each season covering the same period.
Handling Missing Data: After inspecting all the samples, I confirmed that there were no missing values, so no further treatment for missing data was necessary.
After the cleaning process, the dataset was reduced to three variables: Time, Total Deaths, and Season, as shown in Table 2.

```{r}
#| warning: false
#| message: false
#| echo: false
#| #| label: Sample of the cleaned data
#| tbl-cap: "Sample of the cleaned data"
kable(head(analysis_data))
```

## Statistical summary of the cleaned data
Table 3 provides some basic statistics about the dataset. The project spans a total of 204 months. During this period, the highest number of deaths recorded in a single month was 19, while the lowest was 0. The average number of deaths per month is 3.848.
```{r}
#| warning: false
#| message: false
#| echo: false
#| label: Basic statistics of the cleaned data
#| tbl-cap: "Basic statistics of the cleaned data"
kable(summary(analysis_data[, c("year_month", "total_decedents")]))
```

Figure 1 illustrates the variation in the number of deaths per month over time. It is clear that the trend is upward, indicating an increase in the number of deaths over the observed period.

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: Total Decedents Over time
#| fig-cap: "Total Decedents Over time"

# Line plot for total_decedents over time
ggplot(analysis_data, aes(x = as.Date(paste(year_month, "-01", sep="")), y = total_decedents)) +
  geom_line() +
  labs(title = "Total Decedents Over Time", x = "Date", y = "Total Decedents") +
  theme_minimal()
```

# Result

Figure 2 presents the fitted results of a simple linear regression model. The blue line represents the predicted data, while the points indicate the actual observed data. It is evident that the number of deaths shows a clear upward trend. The slope of the line suggests that the death count is increasing at an approximate rate of 1.009105 per week.

```{r}
#| warning: false
#| message: false
#| echo: false
#| include: false
# 将 year_month 转换为日期格式，假设 year_month 是 YYYY-MM 格式的字符型
analysis_data$year_month_date <- as.Date(paste0(analysis_data$year_month, "-01"))

# 设置起始日期为 2007-01-01
start_date <- as.Date("2007-01-01")

# 计算每个日期与起始日期之间的时间差（以天为单位）
analysis_data$days_since_start <- as.numeric(difftime(analysis_data$year_month_date, start_date, units = "weeks"))

# 使用计算出的时间差进行回归分析
m <- lm(total_decedents ~ days_since_start, data = analysis_data)
```

```{r, results='hide'}
#| warning: false
#| message: false
#| echo: false
#| label: Simple linear regression
#| fig-cap: "Simple linear regression"

ggplot(analysis_data, aes(x = ym(year_month), y = total_decedents)) +
  geom_point() +  # 绘制数据点
  geom_line(aes(y = fitted(m)), col = "blue", size = 1)
  labs(title = "Total Decedents vs Year/Month",
       x = "Time",
       y = "Total Decedents")
```
Figure 3 displays the total number of deaths per year, along with their respective proportions. It is important to note that I removed the data for 2007 and 2024 because they do not represent complete years. From the chart, it is evident that the number of deaths between 2020 and 2023 was significantly higher, with each of these years accounting for over 10% of the total deaths during the 15-year period. The peak occurred in 2021, reaching as high as 17.8%.

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: Total Decedents per Year
#| fig-cap: "Total Decedents per Year"

# 将 year_month 转换为年份
analysis_data$year <- substr(analysis_data$year_month, 1, 4)

# 按年份汇总总死亡人数
yearly_deaths <- aggregate(total_decedents ~ year, data = analysis_data, sum)

filtered_data <- yearly_deaths[!(yearly_deaths$year %in% c(2007, 2024)), ]

# 计算每年的总死亡人数占比（百分比）
filtered_data$total_decedents_pct <- filtered_data$total_decedents / sum(filtered_data$total_decedents) * 100

# 绘制带有百分比标签的柱状图
ggplot(filtered_data, aes(x = year, y = total_decedents)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  geom_text(aes(label = paste0(round(total_decedents_pct, 1), "%")), 
            vjust = -0.5, size = 3) +  # 添加百分比标签
  labs(title = "Total Decedents per Year", x = "Year", y = "Total Decedents") +
  theme_minimal()

```

Additionally, Figure 4 shows the average number of deaths per season over the past few years. The proportions for spring, autumn, and summer are relatively similar, ranging between 23.6% and 23.8%. However, winter stands out with a significantly higher proportion, reaching 29%, notably surpassing the other three seasons.

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: Average Decedents by Season
#| fig-cap: "Average Decedents by Season"

seasonal_data <- analysis_data %>%
  group_by(season) %>%
  summarize(mean_decedents = mean(total_decedents),
            total_decedents = sum(total_decedents))

# 计算每个季节的百分比
seasonal_data$total_decedents_pct <- seasonal_data$total_decedents / sum(seasonal_data$total_decedents) * 100

# 绘制带有百分比的柱状图
ggplot(seasonal_data, aes(x = season, y = mean_decedents, fill = season)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = paste0(round(total_decedents_pct, 1), "%")), 
            vjust = -0.5, size = 3) +  # 添加百分比标签
  labs(title = "Average Decedents by Season", 
       x = "Season", 
       y = "Average Decedents") +
  theme_minimal()
```

# Discussion
\newpage

\appendix

# Appendix {-}
\newpage

# References

### 2024-09-23 scripts/04-model.R

@@ -10,28 +10,17 @@
 
 #### Workspace setup ####
 library(tidyverse)
-library(rstanarm)
 
+getwd()
 #### Read data ####
 analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
 
-### Model data ####
-first_model <-
-  stan_glm(
-    formula = flying_time ~ length + width,
-    data = analysis_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
-  )
 
+
+
+
+### Model data ####
 
 #### Save model ####
-saveRDS(
-  first_model,
-  file = "models/first_model.rds"
-)
 
 

### 2024-09-24 paper/paper.qmd
paper updated
1. second draft of paper and corresponding references updated.
2. sketches updated

@@
   - Haowei Fan
-thanks: "Data are available at: https://open.toronto.ca/dataset/deaths-of-shelter-residents/"
 date: today
@@
 
-In 2023, as the COVID-19 pandemic receded, the Toronto government announced plans to phase out the shelter hotels that had been providing temporary accommodations for homeless individuals.
-By the end of 2024, 24 shelter hotels will no longer be available due to expiring contracts, exacerbating an already overcrowded shelter system.
+In 2023, as the COVID-19 pandemic receded, the Toronto government announced plans to phase out the shelter hotels that had been providing temporary accommodations for homeless individuals (@NoHotel).
+By the end of 2024, 24 shelter hotels will no longer be available due to expiring contracts, exacerbating an already overcrowded shelter system (@NoHotel).
 The loss of these temporary accommodations poses significant risks not only to the safety and stability of the community but also to the well-being of those who rely on shelters for survival.
@@
 This highlights the urgent need to understand the trends and underlying factors affecting shelter residents, especially in terms of their mortality numbers.
-While many sociological studies have explored the growing number of shelter residents, fewer have examined shelter resident mortality rates in detail.
+
+While many sociological studies have explored the growing number of shelter residents.
+For example, Maureen Crane and Louise Joly have studied the growing number of shelter residents and the issue of aging in Western countries (@Crane_Joly_2014).
+Similarly, in their article, John M.
+Quigley, Steven Raphael, and Eugene Smolensky pointed out that the number of shelter residents in the United States is increasing, along with the underlying causes contributing to this rise (@AmericanHomeless).
+Fewer have examined shelter resident mortality rates in detail.
 Given the challenges of accurately counting the transient and often untracked shelter population, studying actual death counts offers a more reliable and feasible method of understanding the impact of systemic issues.
-Mortality data, being both concrete and verifiable, can better inform policy decisions than estimates of shelter population numbers, which are often difficult to obtain and track.
+Mortality data, being both concrete and verifiable, can better inform policy decisions than estimates of shelter population numbers, which are often difficult to obtain and track (@LackAppraoch).
+
 This paper focuses on analyzing the number of deaths among shelter residents in Toronto from March 2007 to August 2024.
@@
 Using simple linear regression, this study estimates the rate at which the number of deaths has increased over time.
-It also examines anomalies in the data by comparing yearly death rates, and assesses seasonal differences by analyzing average death rates across the four seasons.
+This project also examines anomalies in the data by comparing yearly death numbers, and assesses seasonal differences by analyzing average death rates across the four seasons.
 This analysis provides valuable insights into the hardships faced by shelter residents, especially during specific times of the year.
-This study finds three key trends: 1) the number of deaths among shelter residents has been rising over time; 2) there was a noticeable increase in deaths between 2020 and 2023; and 3) while deaths during spring, summer, and fall remained relatively consistent, winter saw significantly higher mortality rates.
+
+This study finds three key trends: 1) the number of deaths among shelter residents in Toronto has been rising over time; 2) there was a noticeable increase in deaths between 2020 and 2023; and 3) while deaths during spring, summer, and fall remained relatively consistent, winter saw significantly higher mortality rates.
 Understanding these trends is critical because shelter resident deaths serve as a key indicator of broader social problems, such as gaps in welfare systems, housing shortages, and insufficient healthcare services.
 These findings provide essential data for governments and welfare organizations to develop more targeted interventions based on the needs of shelter residents throughout the year.
+
 The remainder of this paper is structured as follows: Section 2 discusses the data sources used in this project and details the data cleaning process, followed by a visualization of the dataset's key distribution and statistics.
@@
 
-The dataset I am using is titled "Deaths of Shelter Residents", sourced from OpenToronto.
+The dataset I am using is titled "Deaths of Shelter Residents", sourced from OpenToronto (@Opendatatoronto).
 This dataset records the number of deaths among shelter residents in the Toronto area, starting from January 2007 on a monthly basis.
-It also includes the gender of the deceased, categorized into three sections: male, female, and other gender, which includes Transgender, Non-binary, and Two-Spirit.
+It also includes the gender of the deceased, categorized into three sections: male, female, and other gender, which includes Transgender, Non-binary, and Two-Spirit (@Opendatatoronto).
 The dataset is continuously updated.
-Table 1 contains descriptions of these variables.
+@tbl-column contains descriptions of these variables.
 
@@
 #| echo: false
-#| label: Column descriptions of the original data
+#| label: tbl-column
 #| tbl-cap: "Column descriptions of the original data"
@@
 
-In addition to this dataset, OpenToronto also offers a similar dataset called "Deaths of People Experiencing Homelessness", which includes four variables: id, year, month, and number of deaths.
+In addition to this dataset, OpenToronto also offers a similar dataset called "Deaths of People Experiencing Homelessness", which includes four variables: id, year, month, and number of deaths (@Opendatatoronto).
 However, I chose not to use this dataset due to its limited data coverage, spanning only from May 2007 to August 2023.
-In comparison, the "Deaths of Shelter Residents" dataset provides more extensive data, allowing for greater accuracy and a narrower confidence interval in the results.
+In comparison, the "Deaths of Shelter Residents" dataset provides more extensive data, allowing for greater accuracy in the results.
 
@@
 
-In the data cleaning process, the following steps were conducted using Rcode(@citeR), along with the packages pandas, tidyverse, ggplot2, and dplyr:
+During the data processing phase, tasks such as downloading, cleaning, simulating, analyzing, and code testing were all performed using R code (@citeR).
+Additionally, the project utilized packages such as Tidyverse (@Tidyverse), Dplyr (@Dplyr), Lubridate (@Lubridate), and Knitr (@Knitr) for data handling.
+These tools were instrumental in performing linear regression, creating visualizations, generating tables, and formatting data for analysis.
 
-Combining Year and Month: The variables for year and month were merged, and the month was converted to a numerical format to facilitate time series analysis.
-Creating a New Variable for Seasons: A new variable was created to categorize each month into one of the four seasons: March, April, and May were classified as spring; June, July, and August as summer; September, October, and November as autumn; and December, January, and February as winter.
-This was done to enable seasonal analysis.
-Removing Unnecessary Variables: All variables unrelated to this project, such as id and the three gender categories, were removed to simplify the dataset.
-Ensuring Equal Seasonal Data: To ensure that each season had the same length of time for comparison, I removed some samples from the beginning and end of the dataset, so that the total number of samples became a multiple of four, with each season covering the same period.
-Handling Missing Data: After inspecting all the samples, I confirmed that there were no missing values, so no further treatment for missing data was necessary.
-After the cleaning process, the dataset was reduced to three variables: Time, Total Deaths, and Season, as shown in Table 2.
+The data cleaning process involved several key steps, each designed to improve the dataset's suitability for analysis.
+First, the year and month variables were combined, with the month converted to numerical format to facilitate time series analysis.
+A new variable was then created to classify each month into one of the four seasons: March, April, and May as spring; June, July, and August as summer; September, October, and November as autumn; and December, January, and February as winter, enabling seasonal analysis.
+To streamline the dataset, unnecessary variables, such as ID and the three gender categories, were removed.
+Additionally, to ensure that the seasonal analysis of average shelter resident deaths covered equal time intervals for each season, certain samples from the dataset's beginning and end were removed, resulting in a total number of samples divisible by four.
+Finally, after reviewing the dataset, it was confirmed that there were no missing values, so no further treatment was necessary.
+Upon completion, the cleaned dataset was reduced to three variables: Time, Total Deaths, and Season, as outlined in @tbl-sample.
 
@@
 #| echo: false
-#| #| label: Sample of the cleaned data
+#| label: tbl-sample
 #| tbl-cap: "Sample of the cleaned data"
@@
 ## Statistical summary of the cleaned data
-Table 3 provides some basic statistics about the dataset.
+@tbl-sumarry provides some basic statistics about the dataset.
 The project spans a total of 204 months.
@@
 The average number of deaths per month is 3.848.
+Additionally, we can infer that the data is somewhat skewed, as the highest number of deaths in a single month is 19, while the average and median number of deaths are 3 and 3.83, respectively.
 ```{r}
@@
 #| echo: false
-#| label: Basic statistics of the cleaned data
+#| label: tbl-sumarry
 #| tbl-cap: "Basic statistics of the cleaned data"
@@
 
-Figure 1 illustrates the variation in the number of deaths per month over time. It is clear that the trend is upward, indicating an increase in the number of deaths over the observed period.
+@fig-deathovertime depicts the monthly variation in the number of deaths over time.
+The overall trend reveals a significant increase in the total number of shelter resident deaths in Toronto, particularly after 2015.
+Prior to 2015, the number of deaths remained relatively stable, rarely exceeding 5 per month.
+However, from 2015 onwards, there is not only a noticeable upward trend in deaths but also increased volatility.
+Spikes in the data become more frequent, with several months recording over 10 deaths, and some months reaching as high as 15 or more.
 
@@
 #| echo: false
-#| label: Total Decedents Over time
+#| label: fig-deathovertime
 #| fig-cap: "Total Decedents Over time"
@@
 # Result
-
-Figure 2 presents the fitted results of a simple linear regression model.
-The blue line represents the predicted data, while the points indicate the actual observed data.
-It is evident that the number of deaths shows a clear upward trend.
-The slope of the line suggests that the death count is increasing at an approximate rate of 1.009105 per week.
 
@@
 ```
-
+I used the total number of shelter resident deaths per month as the target variable to be predicted, with time and season as predictors, and performed a linear regression analysis.
+Since time and season are evidently highly correlated, I decided to retain only time as the final predictor for the model.
+After completing the simple linear regression, I obtained a positive slope, represented by a straight line (@fig-slr), indicating that the number of shelter resident deaths in Toronto has increased over time.
+A comparison with the actual data revealed that the real values closely align with the predictions from the SLR model, demonstrating the reliability of my model.
 ```{r, results='hide'}
@@
 #| echo: false
-#| label: Simple linear regression
+#| label: fig-slr
 #| fig-cap: "Simple linear regression"
@@
 ```
+
+####################
 Figure 3 displays the total number of deaths per year, along with their respective proportions.

### 2024-09-24 paper/paper.qmd
Formatting erros corrected.
Some formatting erros corrected.

@@
 #| echo: false
+#### Workspace setup ####
 library(tidyverse)
 library(knitr)
+#### Read data ####
 analysis_data <- read_csv("/Users/fhw0912/Desktop/Deaths-of-Shelter-Residents-in-Toronto/data/analysis_data/analysis_data.csv")
@@
 
+# Build a table to display columns and thier description.
 column_description <- tibble(
@@
 
+# Draw the table.
 kable(column_description, col.names = c("Column", "Description"))
@@
 #| tbl-cap: "Sample of the cleaned data"
+
+# Display few of samples.
 kable(head(analysis_data))
@@
 #| tbl-cap: "Basic statistics of the cleaned data"
+
+# Showing the summary of cleaned data.
 kable(summary(analysis_data[, c("year_month", "total_decedents")]))
@@
 #| include: false
-# 将 year_month 转换为日期格式，假设 year_month 是 YYYY-MM 格式的字符型
+# Convert year_month to a date format, assuming year_month is a character string in the format YYYY-MM
 analysis_data$year_month_date <- as.Date(paste0(analysis_data$year_month, "-01"))
 
-# 设置起始日期为 2007-01-01
+# Set the start date to 2007-01-01
 start_date <- as.Date("2007-01-01")
 
-# 计算每个日期与起始日期之间的时间差（以天为单位）
+# Calculate the time difference (in weeks) between each date and the start date
 analysis_data$days_since_start <- as.numeric(difftime(analysis_data$year_month_date, start_date, units = "weeks"))
 
-# 使用计算出的时间差进行回归分析
+# Perform a regression analysis using the calculated time difference
 m <- lm(total_decedents ~ days_since_start, data = analysis_data)
@@
 
+# Scatter plot and fitted line
 ggplot(analysis_data, aes(x = ym(year_month), y = total_decedents)) +
-  geom_point() +  # 绘制数据点
+  geom_point() +
   geom_line(aes(y = fitted(m)), col = "blue", size = 1)
@@
 
-####################
-Figure 3 displays the total number of deaths per year, along with their respective proportions.
-It is important to note that I removed the data for 2007 and 2024 because they do not represent complete years.
-From the chart, it is evident that the number of deaths between 2020 and 2023 was significantly higher, with each of these years accounting for over 10% of the total deaths during the 15-year period.
-The peak occurred in 2021, reaching as high as 17.8%.
+@fig-yeardeath displays the total number of shelter resident deaths per year, along with their respective proportions.
+It is important to note that I excluded the data for 2007 and 2024, as these are incomplete years and including them could mislead the reader.
+From the chart, it is clear that the number of deaths between 2020 and 2023 was significantly higher, with each of these years accounting for over 10% of the total deaths during the 15-year period.
+The peak occurred in 2021, where the number of deaths exceeded 125, representing 17.8% of the total.
 
@@
 #| echo: false
-#| label: Total Decedents per Year
+#| label: fig-yeardeath
 #| fig-cap: "Total Decedents per Year"
 
-# 将 year_month 转换为年份
+# Convert year_month to year
 analysis_data$year <- substr(analysis_data$year_month, 1, 4)
 
-# 按年份汇总总死亡人数
+# Aggregate total deaths by year
 yearly_deaths <- aggregate(total_decedents ~ year, data = analysis_data, sum)
 
+# Filter out data for the years 2007 and 2024
 filtered_data <- yearly_deaths[!(yearly_deaths$year %in% c(2007, 2024)), ]
 
-# 计算每年的总死亡人数占比（百分比）
+# Calculate the percentage of total deaths per year
 filtered_data$total_decedents_pct <- filtered_data$total_decedents / sum(filtered_data$total_decedents) * 100
 
-# 绘制带有百分比标签的柱状图
+# Create a bar chart with percentage labels
 ggplot(filtered_data, aes(x = year, y = total_decedents)) +
@@
   geom_text(aes(label = paste0(round(total_decedents_pct, 1), "%")), 
-            vjust = -0.5, size = 3) +  # 添加百分比标签
+            vjust = -0.5, size = 3) +  # Add percentage labels
   labs(title = "Total Decedents per Year", x = "Year", y = "Total Decedents") +
@@
 
-Additionally, Figure 4 shows the average number of deaths per season over the past few years.
+Additionally, @fig-seasondeath shows the average number of deaths per season over the past few years.
 The proportions for spring, autumn, and summer are relatively similar, ranging between 23.6% and 23.8%.
@@
 #| echo: false
-#| label: Average Decedents by Season
+#| label: fig-seasondeath
 #| fig-cap: "Average Decedents by Season"
 
+# Grouping data by season
 seasonal_data <- analysis_data %>%
@@
 
-# 计算每个季节的百分比
+# Calculate the percentage of total deaths for each season
 seasonal_data$total_decedents_pct <- seasonal_data$total_decedents / sum(seasonal_data$total_decedents) * 100
 
-# 绘制带有百分比的柱状图
+# Create a bar chart with percentage labels
 ggplot(seasonal_data, aes(x = season, y = mean_decedents, fill = season)) +
@@
   geom_text(aes(label = paste0(round(total_decedents_pct, 1), "%")), 
-            vjust = -0.5, size = 3) +  # 添加百分比标签
+            vjust = -0.5, size = 3) +  # Add percentage labels
   labs(title = "Average Decedents by Season", 

### 2024-09-24 paper/paper.qmd
README and LLM Claims Done
README and LLM Claims are finished.

@@
 bibliography: references.bib
+
 ---
@@
 #### Read data ####
-analysis_data <- read_csv("/Users/fhw0912/Desktop/Deaths-of-Shelter-Residents-in-Toronto/data/analysis_data/analysis_data.csv")
+analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
 ```
@@
 
-The remainder of this paper is structured as follows: Section 2 discusses the data sources used in this project and details the data cleaning process, followed by a visualization of the dataset's key distribution and statistics.
-Section 3 outlines the statistical methods used and presents the findings in detail.
-Section 4 addresses the limitations of the study and offers suggestions for future research.
+The remainder of this paper is structured as follows: @sec-data discusses the data sources used in this project and details the data cleaning process, followed by a visualization of the dataset's key distribution and statistics.
+@sec-result outlines the statistical methods used and presents the findings in detail.
+@sec-discussion addresses the limitations of the study and offers suggestions for future research.
 
@@
 
-# Result
+# Result {#sec-result}
 
@@
 
-# Discussion
+# Discussion {#sec-discussion}
+
 \newpage

### 2024-09-24 scripts/04-model.R
useless files deleted, skeches uploaded
1. useless files deleted,

2. skeches uploaded

@@ -1,26 +1,0 @@
-#### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
-
-#### Workspace setup ####
-library(tidyverse)
-
-getwd()
-#### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
-
-
-
-
-
-### Model data ####
-
-#### Save model ####
-
-

### 2024-09-24 paper/paper.qmd
discussion section of paper added
discussion section of paper added

@@
 
+## Limitations
+This study acknowledges several limitations.
+First, although the number of shelter resident deaths is more reliably recorded than the total shelter resident population, due to the unstructured nature of this group, the conclusions drawn about death trends may not fully reflect the challenges faced by the entire population.
+For instance, an increase in the total shelter population with a stable mortality rate would still result in a rise in deaths, which may not indicate worsening conditions.
+Second, the project utilizes a relatively small sample of data (204 months), and therefore, the results from the linear regression model may be subject to bias.
+However, this dataset represents the most comprehensive data available from government sources.
+Third, when analyzing seasonal variations, the sharp rise in death counts after 2020 introduces potential bias.
+It is possible that there were no seasonal differences before 2020, but new patterns emerged thereafter, which may have skewed the overall average and affected the seasonal analysis.
+
+## Suggestions for future research
+There are several avenues for future research that could build upon the findings of this study.
+First, as noted in the limitations, this research focuses on the number of shelter resident deaths, which may not fully reflect the conditions of the entire shelter population.
+Future studies examining the relationship between the total shelter population and the number of deaths could provide a more comprehensive understanding of the current situation.
+Second, while seasonal analysis in this study may have been influenced by recent years’ data, it is undeniable that shelter residents face heightened risks during the winter months at least in recent years.
+Research into the specific reasons behind increased winter threats, and an exploration of the actual needs of shelter residents during this period, would provide governments and welfare organizations with clearer insights to improve policies and intervention strategies.
+Finally, this research highlights the surge in death counts after 2020, but the underlying causes behind this spike should be further investigated by future scholars to prevent similar occurrences in the future.
+
 \newpage

### 2024-09-25 paper/paper.qmd
Bugs fixed, Paper Done, Formatted
1. Bugs in test file fixed.
2. Paper finished.
3. Graphs, tables modified.
4. and other minor problem fixed

@@
 date: today
+thanks: "Code and data are available at: https://github.com/HaoweiFan0912/Deaths-of-Shelter-Residents-in-Toronto.git"
 date-format: long
@@
 library(knitr)
+library(dplyr)
+set.seed(912)
 #### Read data ####
@@
 Fewer have examined shelter resident mortality rates in detail.
-Given the challenges of accurately counting the transient and often untracked shelter population, studying actual death counts offers a more reliable and feasible method of understanding the impact of systemic issues.
-Mortality data, being both concrete and verifiable, can better inform policy decisions than estimates of shelter population numbers, which are often difficult to obtain and track (@LackAppraoch).
+Given the challenges of accurately tracking the transient shelter population, studying actual death counts—concrete and verifiable data—offers a more reliable and feasible method for understanding changes in the needs of shelter residents, providing better insights for policy decisions than difficult-to-obtain population estimates (@LackAppraoch).
 
-This paper focuses on analyzing the number of deaths among shelter residents in Toronto from March 2007 to August 2024.
-The aim is to identify both long-term trends and seasonal variations in shelter resident deaths.
-Using simple linear regression, this study estimates the rate at which the number of deaths has increased over time.
-This project also examines anomalies in the data by comparing yearly death numbers, and assesses seasonal differences by analyzing average death rates across the four seasons.
-This analysis provides valuable insights into the hardships faced by shelter residents, especially during specific times of the year.
-
-This study finds three key trends: 1) the number of deaths among shelter residents in Toronto has been rising over time; 2) there was a noticeable increase in deaths between 2020 and 2023; and 3) while deaths during spring, summer, and fall remained relatively consistent, winter saw significantly higher mortality rates.
-Understanding these trends is critical because shelter resident deaths serve as a key indicator of broader social problems, such as gaps in welfare systems, housing shortages, and insufficient healthcare services.
-These findings provide essential data for governments and welfare organizations to develop more targeted interventions based on the needs of shelter residents throughout the year.
+This paper analyzes the number of deaths among shelter residents in Toronto from March 2007 to August 2024, aiming to identify long-term trends and seasonal variations.
+Using simple linear regression, the study estimates the rate of increase in deaths over time, while also examining anomalies by comparing yearly death numbers and assessing seasonal differences through average death rates across the four seasons.
+The findings reveal three key trends: 1) a steady rise in the number of shelter resident deaths over time; 2) a noticeable spike in deaths between 2020 and 2023; and 3) significantly higher mortality rates during winter compared to the relatively consistent rates in spring, summer, and fall.
+These insights provide critical data for governments and welfare organizations to develop more targeted interventions based on the needs of shelter residents throughout times.
 
@@
 
-The dataset I am using is titled "Deaths of Shelter Residents", sourced from OpenToronto (@Opendatatoronto).
-This dataset records the number of deaths among shelter residents in the Toronto area, starting from January 2007 on a monthly basis.
-It also includes the gender of the deceased, categorized into three sections: male, female, and other gender, which includes Transgender, Non-binary, and Two-Spirit (@Opendatatoronto).
+The dataset used in this study is titled "Deaths of Shelter Residents" and is sourced from OpenDataToronto (@Opendatatoronto).
+It was published by Toronto Public Health (TPH) and records the number of deaths among shelter residents in the Toronto area on a monthly basis, starting from January 2007.
+Additionally, the dataset includes the gender distribution of the deceased each month, with gender categorized into three groups: male, female, and other gender, which encompasses Transgender, Non-binary, and Two-Spirit individuals (@Opendatatoronto).
 The dataset is continuously updated.
-@tbl-column contains descriptions of these variables.
+A detailed description of the variables can be found in @tbl-column.
 
@@
 # Draw the table.
-kable(column_description, col.names = c("Column", "Description"))
+kable(column_description, col.names = c("Variables", "Descriptions"))
 ```
 
-In addition to this dataset, OpenToronto also offers a similar dataset called "Deaths of People Experiencing Homelessness", which includes four variables: id, year, month, and number of deaths (@Opendatatoronto).
-However, I chose not to use this dataset due to its limited data coverage, spanning only from May 2007 to August 2023.
-In comparison, the "Deaths of Shelter Residents" dataset provides more extensive data, allowing for greater accuracy in the results.
+In addition to this dataset, OpenDataToronto also provides a similar dataset titled "Deaths of People Experiencing Homelessness," which records monthly deaths of homeless individuals from January 2017 to March 2023 (@Opendatatoronto).
+However, I chose not to use this dataset due to its limited coverage, spanning only from January 2017 to December 2023.
+In comparison, the "Deaths of Shelter Residents" dataset offers a longer time frame, allowing for more comprehensive analysis and greater accuracy in the results (@Opendatatoronto).
 
@@
 Finally, after reviewing the dataset, it was confirmed that there were no missing values, so no further treatment was necessary.
-Upon completion, the cleaned dataset was reduced to three variables: Time, Total Deaths, and Season, as outlined in @tbl-sample.
+Upon completion, the cleaned dataset was reduced to three variables: Time, Season, and Total Deaths, as few samples are outlined in @tbl-sample.
 
@@
 #| label: tbl-sample
-#| tbl-cap: "Sample of the cleaned data"
+#| tbl-cap: "Samples of the cleaned data"
 
 # Display few of samples.
-kable(head(analysis_data))
+kable(head(analysis_data), col.names = c("Time", "Season", "Number of decedents"))
 ```
@@
 #| label: tbl-sumarry
-#| tbl-cap: "Basic statistics of the cleaned data"
+#| tbl-cap: "Summary of the cleaned data"
 
 # Showing the summary of cleaned data.
-kable(summary(analysis_data[, c("year_month", "total_decedents")]))
+kable(summary(analysis_data[, c("year_month", "total_decedents")]),
+      col.names = c("Time", "Number of decedents"))
+```
+
+@fig-decedentdis illustrates the distribution of monthly shelter resident deaths, showing a left-skewed normal distribution.
+This indicates that the majority of months have a death count in the range of 0-5.
+Additionally, there are two outliers, where the death count exceeds 15 in two months, which is an uncommon occurrence.
+
+```{r, fig.pos="H"}
+#| warning: false
+#| message: false
+#| echo: false
+#| label: fig-decedentdis
+#| fig-cap: "Distribution of monthly number of decedents"
+
+# Plotting a histogram for total_decedents
+ggplot(analysis_data, aes(x = total_decedents)) +
+  geom_histogram(binwidth = 1, fill = "steelblue", color = "black", alpha = 0.7) +
+  labs(title = "Distribution of monthly number of decedents", x = "Number of decedents", y = "Frequency") +
+  theme_minimal()+
+  theme(
+    panel.grid.major = element_line(color = "gray80"),  
+    panel.grid.minor = element_blank(),  
+    plot.title = element_text(hjust = 0.5, size = 12),  
+    axis.title.x = element_text(size = 12), 
+    axis.title.y = element_text(size = 12),
+    axis.text.x = element_text(hjust = 0.5,  size = 10),
+    axis.text.y = element_text(hjust = 0,  size = 10),
+  )
 ```
@@
 @fig-deathovertime depicts the monthly variation in the number of deaths over time.
-The overall trend reveals a significant increase in the total number of shelter resident deaths in Toronto, particularly after 2015.
-Prior to 2015, the number of deaths remained relatively stable, rarely exceeding 5 per month.
-However, from 2015 onwards, there is not only a noticeable upward trend in deaths but also increased volatility.
+The overall trend reveals a significant increase in the total number of shelter resident deaths in Toronto, particularly after 2018.
+Prior to 2018, the number of deaths remained relatively stable, rarely exceeding 5 per month.
+However, from 2018 onwards, there is not only a noticeable upward trend in deaths but also increased volatility.
 Spikes in the data become more frequent, with several months recording over 10 deaths, and some months reaching as high as 15 or more.
 
-```{r}
+```{r, fig.pos="H"}
 #| warning: false
@@
 #| label: fig-deathovertime
-#| fig-cap: "Total Decedents Over time"
+#| fig-cap: "Number of decedents over time"
 
-# Line plot for total_decedents over time
+# plot an graph for the trend of number of decedents over time.
 ggplot(analysis_data, aes(x = as.Date(paste(year_month, "-01", sep="")), y = total_decedents)) +
-  geom_line() +
-  labs(title = "Total Decedents Over Time", x = "Date", y = "Total Decedents") +
-  theme_minimal()
+  geom_line(color = "steelblue", size = 0.7) +  
+  labs(title = "Number of decedents over time", 
+       x = "Time", 
+       y = "Number of decedents") +
+  scale_x_date(date_labels = "%Y", date_breaks = "2 years") +  
+  theme_minimal(base_size = 15) +  
+  theme(
+    panel.grid.major = element_line(color = "gray80"),  
+    panel.grid.minor = element_blank(),  
+    plot.title = element_text(hjust = 0.5, size = 12),  
+    axis.title.x = element_text(size = 12), 
+    axis.title.y = element_text(size = 12),
+    axis.text.x = element_text(angle = 30, hjust = 0.5,  size = 10),
+    axis.text.y = element_text(hjust = 0,  size = 10),
+  )
 ```
+## Discussion of measurement
+
+The "Deaths of Shelter Residents" dataset is collected by Toronto Public Health (TPH), a governmental body (@Opendatatoronto).
+Information is sourced from several avenues, including a secure online form where agencies supporting homeless and under-housed populations report deaths (@Opendatatoronto).
+Other sources include the Toronto Shelter & Support Services (TSSS) division and monthly reports from the Toronto Homeless Memorial.
+Any duplicate entries are removed, and data are cross-checked by the Office of the Chief Coroner of Ontario (OCCO) (@Opendatatoronto).
+
+Given that this dataset originates from major institutions in Toronto, its credibility is very high.
+The additional verification by the Office of the Chief Coroner of Ontario (OCCO) further enhances the reliability of the data.
+
+Additionally, collecting data on the deaths of shelter residents is more reliable compared to tracking the entire population.
+First, Toronto is a large metropolitan area, where the presence of witnesses makes data falsification or omission nearly impossible.
+Second, tracking deaths is more dependable than attempting to count the entire population of shelter residents, as this group tends to be less organized, more transient, and harder to communicate with.
+As a result, it is not feasible to accurately account for the entire population in detail (@EndingH).
 
@@
 
-```{r}
-#| warning: false
-#| message: false
-#| echo: false
-#| include: false
-# Convert year_month to a date format, assuming year_month is a character string in the format YYYY-MM
-analysis_data$year_month_date <- as.Date(paste0(analysis_data$year_month, "-01"))
-
-# Set the start date to 2007-01-01
-start_date <- as.Date("2007-01-01")
-
-# Calculate the time difference (in weeks) between each date and the start date
-analysis_data$days_since_start <- as.numeric(difftime(analysis_data$year_month_date, start_date, units = "weeks"))
-
-# Perform a regression analysis using the calculated time difference
-m <- lm(total_decedents ~ days_since_start, data = analysis_data)
-```
 I used the total number of shelter resident deaths per month as the target variable to be predicted, with time and season as predictors, and performed a linear regression analysis.
@@
 A comparison with the actual data revealed that the real values closely align with the predictions from the SLR model, demonstrating the reliability of my model.
-```{r, results='hide'}
+
+```{r, results='hide', fig.pos="H"}
 #| warning: false
@@
 #| label: fig-slr
-#| fig-cap: "Simple linear regression"
+#| fig-cap: "Simple linear regression of number of decedent by time"
 
@@
   geom_point() +
-  geom_line(aes(y = fitted(m)), col = "blue", size = 1)
-  labs(title = "Total Decedents vs Year/Month",
+  geom_smooth(method = "lm", se = FALSE) +
+  labs(title = "Simple linear regression of number of decedent by time",
        x = "Time",
-       y = "Total Decedents")
+       y = "Number of Decedents") +
+  theme(
+    panel.grid.major = element_line(color = "gray80"),  
+    panel.grid.minor = element_blank(),  
+    plot.title = element_text(hjust = 0.5, size = 12),  
+    axis.title.x = element_text(size = 12), 
+    axis.title.y = element_text(size = 12),
+    axis.text.x = element_text(hjust = 0.5,  size = 10),
+    axis.text.y = element_text(hjust = 0,  size = 10),
+  )
 ```
@@
 
-```{r}
+```{r, fig.pos="H"}
 #| warning: false
@@
 #| label: fig-yeardeath
-#| fig-cap: "Total Decedents per Year"
+#| fig-cap: "Number of decedents per Year"
 
@@
   geom_text(aes(label = paste0(round(total_decedents_pct, 1), "%")), 
-            vjust = -0.5, size = 3) +  # Add percentage labels
-  labs(title = "Total Decedents per Year", x = "Year", y = "Total Decedents") +
-  theme_minimal()
+            vjust = -0.5, size = 2.3) +  # Add percentage labels
+  labs(title = "Number of decedents per year", x = "Year", y = "Number of decedents") +
+  theme_minimal()+
+  theme(
+    panel.grid.major = element_line(color = "gray80"),  
+    panel.grid.minor = element_blank(),  
+    plot.title = element_text(hjust = 0.5, size = 12),  
+    axis.title.x = element_text(size = 12), 
+    axis.title.y = element_text(size = 12),
+    axis.text.x = element_text(angle = 30, hjust = 0.5,  size = 10),
+    axis.text.y = element_text(hjust = 0,  size = 10),
 
+)
 ```
@@
 #| label: fig-seasondeath
-#| fig-cap: "Average Decedents by Season"
+#| fig-cap: "Average number of decedents per season"
 
@@
 
+# Rearranging season.
+seasonal_data$season <- factor(seasonal_data$season, levels = c("Spring", "Summer", "Autumn", "Winter"))
+seasonal_data_sorted <- seasonal_data %>% arrange(season)
+
 # Create a bar chart with percentage labels
@@
   geom_text(aes(label = paste0(round(total_decedents_pct, 1), "%")), 
-            vjust = -0.5, size = 3) +  # Add percentage labels
-  labs(title = "Average Decedents by Season", 
+            vjust = -0.5, size = 2.5) +  # Add percentage labels
+  labs(title = "Average number of decedents per season", 
        x = "Season", 
-       y = "Average Decedents") +
-  theme_minimal()
+       y = "Average number of decedents") +
+  theme_minimal() +
+  theme(
+    panel.grid.major = element_line(color = "gray80"),  
+    panel.grid.minor = element_blank(),  
+    plot.title = element_text(hjust = 0.5, size = 12),  
+    axis.title.x = element_text(size = 12), 
+    axis.title.y = element_text(size = 12),
+    axis.text.x = element_text(hjust = 0.5,  size = 10),
+    axis.text.y = element_text(hjust = 0,  size = 10),
+  )
 ```

### 2024-09-25 scripts/02-data_cleaning.R

@@ -31,7 +31,5 @@ cleaned_data <-
   filter(ym(year_month) >= ym("2007-03") & ym(year_month) <= ym("2024-02")) |> # delete unwanted rows.
   tidyr::drop_na() # drop rows contains empty unit.
 
-  
-cleaned_data
 #### Save data ####
 write_csv(cleaned_data, "data/analysis_data/analysis_data.csv")

### 2024-09-25 scripts/03-test_data.R

@@ -13,20 +13,20 @@ library(tidyverse)
 #### Test data ####
 
 # select target data
-data <- read_csv("data/analysis_data/simulated_data.csv")
+simulated_data <- read_csv("data/analysis_data/simulated_data.csv")
 
 # Test 1: Ensure year_month format is correct (YYYY-MM)
-all(grepl("^\\d{4}-\\d{2}$", data$year_month))
+all(grepl("^\\d{4}-\\d{2}$", simulated_data$year_month))
 
 
 # Test 2: Ensure year_month is in the right period.
-all(ym(data$year_month) >= ym("2007-03") 
-    & ym(data$year_month) <= ym("2024-02"))
+all(ym(simulated_data$year_month) >= ym("2007-03") 
+    & ym(simulated_data$year_month) <= ym("2024-02"))
 
 
 # Test 3: Ensure all seasons are assigned correctly
 valid_seasons <- c("Spring", "Summer", "Autumn", "Winter")
-all(data$season %in% valid_seasons)
+all(simulated_data$season %in% valid_seasons)
 
 
 # Test 4: the actual season matches the expected season
@@ -43,32 +43,33 @@ get_season <- function(month) {
   }
 }
 # Test to check if the 'season' column matches the correct season
-correct_season <- tibble(correct_season = sapply(data$month, get_season))
-all(correct_season$correct_season == data$season)
+month_from_data <- as.numeric(substr(simulated_data$year_month, 6, 7))
+correct_season <- tibble(correct_season = sapply(month_from_data, get_season))
+all(correct_season$correct_season == simulated_data$season)
 
 
 # Test 5: Check for valid total decedents
-all(data$total_decedents >= 0 & # is non-negative?
-      data$total_decedents == as.integer(data$total_decedents)) # is integer?
+all(simulated_data$total_decedents >= 0 & # is non-negative?
+      simulated_data$total_decedents == as.integer(simulated_data$total_decedents)) # is integer?
 
 # now I will test the actual data.
-data <- read_csv("data/analysis_data/analysis_data.csv")
+analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
 
-# Test 7: Ensure year_month format is correct (YYYY-MM)
-all(grepl("^\\d{4}-\\d{2}$", data$year_month))
-
-
-# Test 8: Ensure year_month is in the right period.
-all(ym(data$year_month) >= ym("2007-03") 
-    & ym(data$year_month) <= ym("2024-02"))
[…]

### 2024-09-25 paper/paper.qmd
styled used for formatting
styled used for formatting

@@
 kable(summary(analysis_data[, c("year_month", "total_decedents")]),
-      col.names = c("Time", "Number of decedents"))
+  col.names = c("Time", "Number of decedents")
+)
 ```
@@
   labs(title = "Distribution of monthly number of decedents", x = "Number of decedents", y = "Frequency") +
-  theme_minimal()+
+  theme_minimal() +
   theme(
@@
 # plot an graph for the trend of number of decedents over time.
-ggplot(analysis_data, aes(x = as.Date(paste(year_month, "-01", sep="")), y = total_decedents)) +
+ggplot(analysis_data, aes(x = as.Date(paste(year_month, "-01", sep = "")), y = total_decedents)) +
   geom_line(color = "steelblue", size = 0.7) +
-  labs(title = "Number of decedents over time", 
+  labs(
+    title = "Number of decedents over time",
     x = "Time",
-       y = "Number of decedents") +
+    y = "Number of decedents"
+  ) +
   scale_x_date(date_labels = "%Y", date_breaks = "2 years") +
@@
   geom_smooth(method = "lm", se = FALSE) +
-  labs(title = "Simple linear regression of number of decedent by time",
+  labs(
+    title = "Simple linear regression of number of decedent by time",
     x = "Time",
-       y = "Number of Decedents") +
+    y = "Number of Decedents"
+  ) +
   theme(
@@
   geom_text(aes(label = paste0(round(total_decedents_pct, 1), "%")),
-            vjust = -0.5, size = 2.3) +  # Add percentage labels
+    vjust = -0.5, size = 2.3
+  ) + # Add percentage labels
   labs(title = "Number of decedents per year", x = "Year", y = "Number of decedents") +
-  theme_minimal()+
+  theme_minimal() +
   theme(
@@
     axis.text.y = element_text(hjust = 0, size = 10),
-
   )
@@
   group_by(season) %>%
-  summarize(mean_decedents = mean(total_decedents),
-            total_decedents = sum(total_decedents))
+  summarize(
+    mean_decedents = mean(total_decedents),
+    total_decedents = sum(total_decedents)
+  )
 
@@
   geom_text(aes(label = paste0(round(total_decedents_pct, 1), "%")),
-            vjust = -0.5, size = 2.5) +  # Add percentage labels
-  labs(title = "Average number of decedents per season", 
+    vjust = -0.5, size = 2.5
+  ) + # Add percentage labels
+  labs(
+    title = "Average number of decedents per season",
     x = "Season",
-       y = "Average number of decedents") +
+    y = "Average number of decedents"
+  ) +
   theme_minimal() +

### 2024-09-25 scripts/00-simulate_data.R

@@ -27,7 +27,7 @@ months <- seq(1, 12)
 seasons <- rep(c("spring", "summer", "autumn", "winter"))
 
 # I assume the number of decedents follows poisson distribution with mean of 8.
-total_decedents <- rpois(length(years)*12, 8)
+total_decedents <- rpois(length(years) * 12, 8)
 
 # Begin to simulate data
 simulated_data <-
@@ -37,7 +37,7 @@ simulated_data <-
     total_decedents = total_decedents
   )
 
-# Add another column to the data called season. 
+# Add another column to the data called season.
 # When month is "Mar", "Apr", or "May", season is Spring.
 # When month is "Jun", "Jul", or "Aug", season is Summer.
 # When month is "Sep", "Oct", or "Nov", season is Autumn.
@@ -47,7 +47,7 @@ simulated_data <- simulated_data %>%
     month %in% c(3, 4, 5) ~ "Spring",
     month %in% c(6, 7, 8) ~ "Summer",
     month %in% c(9, 10, 11) ~ "Autumn",
-    month %in% c(12, 1, 2 ) ~ "Winter"
+    month %in% c(12, 1, 2) ~ "Winter"
   ))
 
 # Format the year_month column.
@@ -61,7 +61,7 @@ simulated_data <- simulated_data %>%
 # Convert year_month to a date using ym (lubridate package)
 simulated_data <- simulated_data %>%
   mutate(year_month = ym(year_month))
-  
+
 # I am only interested in the period from 2007-3 to 2024-2.
 simulated_data <- simulated_data %>%
   filter(year_month >= ym("2007-03") & year_month <= ym("2024-02")) %>%

### 2024-09-25 scripts/01-download_data.R

@@ -14,20 +14,19 @@ library(tidyverse)
 
 #### Download data ####
 
-#Looking at first 200 pages of datasets from "opendatatoronto".
+# Looking at first 200 pages of datasets from "opendatatoronto".
 list_packages <- list_packages(200)
 
-#Searching for the dataset called "Deaths of Shelter Residents" which I am interested in.
+# Searching for the dataset called "Deaths of Shelter Residents" which I am interested in.
 list_deathShelter <- search_packages("Deaths of Shelter Residents")
 
-#Get the id of the dataset.
+# Get the id of the dataset.
 list_deathShelter_resources <- list_deathShelter %>%
   list_package_resources()
 
-#Download the data by its' id
-deaths_of_shelter_residents <- list_deathShelter_resources[3,"id"] %>%
+# Download the data by its' id
+deaths_of_shelter_residents <- list_deathShelter_resources[3, "id"] %>%
   get_resource()
 
 #### Save data ####
-write_csv(deaths_of_shelter_residents, "data/raw_data/raw_data.csv") 
-
+write_csv(deaths_of_shelter_residents, "data/raw_data/raw_data.csv")

### 2024-09-25 scripts/02-data_cleaning.R

@@ -17,9 +17,11 @@ cleaned_data <-
   raw_data |>
   janitor::clean_names() |> # format column names
   select(-x_id, -transgender_non_binary_two_spirit, -male, -female) |> # delete unless columns
-  mutate(month_numeric = sprintf("%02d", match(month, month.abb)),  # Convert month abbreviation to numeric
-         year_month = paste(year, month_numeric, sep = "-")) %>% # New column contains year and month in the format XXXX_XX
-  select(-year, -month) |>   # Keep only the new 'year_month' column
+  mutate(
+    month_numeric = sprintf("%02d", match(month, month.abb)), # Convert month abbreviation to numeric
+    year_month = paste(year, month_numeric, sep = "-")
+  ) %>% # New column contains year and month in the format XXXX_XX
+  select(-year, -month) |> # Keep only the new 'year_month' column
   mutate(season = case_when(
     month_numeric %in% c("03", "04", "05") ~ "Spring", # match months to actual season
     month_numeric %in% c("06", "07", "08") ~ "Summer",

### 2024-09-25 scripts/03-test_data.R

@@ -20,8 +20,8 @@ all(grepl("^\\d{4}-\\d{2}$", simulated_data$year_month))
 
 
 # Test 2: Ensure year_month is in the right period.
-all(ym(simulated_data$year_month) >= ym("2007-03") 
-    & ym(simulated_data$year_month) <= ym("2024-02"))
+all(ym(simulated_data$year_month) >= ym("2007-03") &
+  ym(simulated_data$year_month) <= ym("2024-02"))
 
 
 # Test 3: Ensure all seasons are assigned correctly
@@ -50,7 +50,7 @@ all(correct_season$correct_season == simulated_data$season)
 
 # Test 5: Check for valid total decedents
 all(simulated_data$total_decedents >= 0 & # is non-negative?
-      simulated_data$total_decedents == as.integer(simulated_data$total_decedents)) # is integer?
+  simulated_data$total_decedents == as.integer(simulated_data$total_decedents)) # is integer?
 
 # now I will test the actual data.
 analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
@@ -60,8 +60,8 @@ all(grepl("^\\d{4}-\\d{2}$", analysis_data$year_month))
 
 
 # Test 7: Ensure year_month is in the right period.
-all(ym(analysis_data$year_month) >= ym("2007-03") 
-    & ym(analysis_data$year_month) <= ym("2024-02"))
+all(ym(analysis_data$year_month) >= ym("2007-03") &
+  ym(analysis_data$year_month) <= ym("2024-02"))
 
 
 # Test 8: Ensure all seasons are assigned correctly
@@ -90,8 +90,4 @@ all(correct_season$correct_season == analysis_data$season)
 
 # Test 10: Check for valid total decedents
 all(analysis_data$total_decedents >= 0 & # is non-negative?
-      analysis_data$total_decedents == as.integer(analysis_data$total_decedents)) # is integer?
-
-
-
-
+  analysis_data$total_decedents == as.integer(analysis_data$total_decedents)) # is integer?

### 2024-09-25 paper/paper.qmd
Minor mistakes fixed
minor formatting mistakes fixed

@@
 bibliography: references.bib
-
+toc: true
 ---
@@
 During the data processing phase, tasks such as downloading, cleaning, simulating, analyzing, and code testing were all performed using R code (@citeR).
-Additionally, the project utilized packages such as Tidyverse (@Tidyverse), Dplyr (@Dplyr), Lubridate (@Lubridate), and Knitr (@Knitr) for data handling.
+Additionally, the project utilized packages such as Tidyverse (@Tidyverse), Styler (@Styler), Dplyr (@Dplyr), Lubridate (@Lubridate), and Knitr (@Knitr) for data handling.
 These tools were instrumental in performing linear regression, creating visualizations, generating tables, and formatting data for analysis.
@@
 Additionally, we can infer that the data is somewhat skewed, as the highest number of deaths in a single month is 19, while the average and median number of deaths are 3 and 3.83, respectively.
+
+\newpage
+
 ```{r}
@@
 Since time and season are evidently highly correlated, I decided to retain only time as the final predictor for the model.
+Additionally, since @fig-decedentdis showed that the total number of decedents follows a normal distribution, it satisfied the conditions for using simple linear regression.
 After completing the simple linear regression, I obtained a positive slope, represented by a straight line (@fig-slr), indicating that the number of shelter resident deaths in Toronto has increased over time.
@@
 
-\appendix
-
-# Appendix {-}
-\newpage
-
 # References

### 2024-09-25 scripts/03-test_data.R

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: Tests codes
+# Purpose: Tests codes.
 # Author: Haowei Fan
 # Date: 22 September 2024
 # Contact: haowei.fan@mail.utoronto.ca

### 2024-09-27 scripts/00-simulate_data.R
Comments on scripts updated
Adding pre-requisities for scripts

@@ -4,7 +4,7 @@
 # Date: 21 September 2024
 # Contact: haowei.fan@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: None
+# Pre-requisites: Understanding the type of data needed for this project.
 # Any other information needed? None
 
 #### Workspace setup ####

### 2024-09-27 scripts/01-download_data.R

@@ -4,7 +4,7 @@
 # Date: 21 September 2024
 # Contact: haowei.fan@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: None
+# Pre-requisites: Being able to access the web: https://open.toronto.ca/dataset/deaths-of-shelter-residents/.
 # Any other information needed? None
 
 

### 2024-09-27 scripts/02-data_cleaning.R

@@ -4,7 +4,7 @@
 # Date: 22 September 2024
 # Contact: haowei.fan@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: None
+# Pre-requisites: The actual data has been downloaded.
 # Any other information needed? None
 
 #### Workspace setup ####

### 2024-09-27 scripts/03-test_data.R

@@ -4,7 +4,7 @@
 # Date: 22 September 2024
 # Contact: haowei.fan@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: None
+# Pre-requisites: Both simulated data and actual raw data are saved in the project folder.
 # Any other information needed? None
 
 #### Workspace setup ####
